Distributional Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain

### 2018-02-22 pkg/disttree/inst/draft/draft.Rnw
small corrections
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2035 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: c352b3b2e7f55bd89de4deb979846abc4d4bbfe2

\subsection{Motivation}
As illustrated in Figure~\ref{Devel_parmod} a wide range of tools has been developed in the field of parametric modeling. 
%The shift from conditional mean regression models to distributional regression models allows for much broader inference. In many applications further features such as the ability to capture non-additive effects or interactions are desirable which demands for more flexibility. 
Especially the shift from conditional mean regression models to distributional regression models has increased the application spectrum of these models as it allows for much broader inference. In some cases further features such as the ability to capture non-additive effects or interactions are desirable which demands for more flexibility. 
Since tree-structured models provide these required features, combining them with parametric models can offer a framework that is able to deal with a wide variety of different modeling situations. In this section the novel methodology which embeds parametric modeling into the idea of tree-structured models will be introduced step by step. Starting with a simple distributional fit in Section~\ref{distfit} will be followed by building a distributional tree in Section~\ref{disttree} and finally a distributional forest in Section~\ref{distforest}.

@@

In our application of precipitation forecasting the response variable, which is the power transformed total precipitation amount within 24 hours, is expected to be normally distributed with mean/location parameter $\mu$ and standard deviation/scale parameter $\sigma$ and left-censored with censoring point 0. (The power transformation is used to remove large parts of the skewness of the data as further explained in Section~\ref{data}.) Therefore, the corresponding log-likelihood function with parameter vector $\theta = (\mu, \sigma)$ is
$$
\ell(\mu, \sigma; Y) = 
\begin{cases}
    \log\left(\frac{1}{\sigma} \cdot \phi\left(\frac{Y - \mu}{\sigma}\right) \right), & \text{if } Y > 0\\
    \vspace{0.1cm}
    \log\left(\Phi\left(\frac{-\mu}{\sigma}\right)\right), & \text{if } Y = 0
\end{cases}
$$
where $\phi$ and $\Phi$ are the probability density function and the distribution function of the standard normal distribution $\mathcal{N}(0,1)$. Next to the censored normal distribution the censored logistic distribution is also often chosen for the modeling of precipitation amounts. In \cite{Gebetsberger+Messner+Mayr:2017} this choice has been made in order to be able to deal with heavy tails which often appear in precipitation data. Another frequent application of probabilistic forecasting in meteorology is the modeling of near surface air temperature. In this case a normal distribution is often selected.

With the specification of the distribution family and the log-likelihood function the task of fitting a distributional model turns into the task of determining the distribution parameter~$\theta$. This is done by the following maximization.
$$
\hat{\theta} = \max_{\theta \in \Theta} \sum_{i=1}^n \ell(\theta; y_i)
$$
where $\{y_i\}_{i=1,\ldots,n}$ are the observations of the response variable $Y$.

In that way, a whole distribution is specified with all its features, including location, scale, and shape. However, fitting one global model to the whole data set might be too much of a generalization and therefore not represent the data and its features in a reasonable way. Moreover, if covariates are available it is desirable to include them into the model as they might provide important additional information. In our application on precipitation data a wide range of covariates are available from a numerical ensemble prediction system considering precipitation, temperature, sea level pressure and other meteorological quantities which might be of interest and have an influence on the predictions.

@@

\subsection{Distributional tree}
\label{disttree}
Given the fit of a global distributional model $D(Y, \bold{\hat{\theta}})$ the idea is to use the information provided by this model together with a set of covariates $\bold{Z} = (Z_1, \ldots, Z_m) \in \mathcal{Z}$ %(from now on denoted by $\bold{Z}$ instead of $\bold{X}$ to indicate that they are considered as split variables only and not as regressors)
%%% <TH> X -> Z is confusing. Either we always write theta = g(X) where
%%% g is now a tree or we explain that theta = g(X)(Z) is what we want.
%%% Here, we don't partition models with X variables, so we should stick to X. </TH>
to decide whether the data set should be split into subgroups or not. In particular, the information provided by the fitted model are the scores 
$$
s(\hat{\theta}, y_i) = \frac{\partial \ell}{\partial \theta}(\hat{\theta}; y_i).
$$ 
These values express how well the model fits the data regarding each parameter. Ideally they should fluctuate randomly around zero, similar to residuals in ordinary least squares estimation, but in this case there is a score value for each pair of estimated parameter $\hat{\theta}_j \in (\hat{\theta}_1, \ldots, \hat{\theta}_k) = \bold{\hat{\theta}}$ and observation $y_j \in \{y_i\}_{i = 1,\ldots, n}$. This enables the model to be sensitive to changes in each of the parameters which is one of the main advantages of distributional trees over other tree models considering only the mean of the distribution.

Statistical tests are applied to check whether there is a significant dependency between the scores and each of the partitioning variables $Z_l \in \bold{Z} = (Z_1, \ldots, Z_m)$. If so, the data set is split into subgroups based on the values of the partitioning variable that shows the highest dependency. The exact way of determining how and where to split varies across different tree algorithms. For distributional trees one possible choice is a permutation test based algorithm which is explained in detail in the appendix. 
After splitting the data the next step is to fit a local distributional model in each of the subgroups. Repeating this procedure of fitting a model, applying statistical tests to evaluate parameter instability and splitting the data set depending on the test results leads to a tree structure. In that way the learning data set is separated into \textit{B} disjoint segments such that
$$
\mathcal{Z} = \dot{\bigcup\limits_{b = 1, \ldots, \textit{B}}} \mathcal{B}_b
$$
where each segment is represented by a terminal node of the tree.

Summed up, the steps of building a distributional tree are:
\begin{enumerate}
\item Specify a distribution with log-likelihood function $\ell(\theta; y)$.
\item Estimate $\bold{\hat{\theta}}$ by solving $\sum_{i=1}^n s(y_i, \hat{\theta}) = 0$.
\item Test for associations or instabilities of the scores $s(\bold{\hat{\theta}}, y_i)$ and each partitioning variable~$Z_l$.
%\item Assess whether the \emph{model scores} are associated with
%    (or change along) any of the available covariates -- e.g.,
%    using parameter instability tests (\emph{strucchange}) or
%    conditional inference (\emph{coin}).
\item Split the sample along the partitioning variable with the strongest association or instability.
    Choose the breakpoint with the highest improvement in log-likelihood.
\item Repeat steps 2--4 recursively in the subgroups
    until some stopping criterion is reached.
    %-- e.g., for significance or sample size.
%\item Choose the variable with the strongest association
%\item Choose the split point which leads to the highest improvement of the model
%\item Split and repeat 2-6 in each node until a stopping criterion is met
\end{enumerate}
The stopping criterion in step 5 can for example be a minimum number of observations in a node or a p-value for the statistical tests applied in step 3.
%%% <TH> we need a little more detail here or a reference to a better
%%% description of mob/ctree OR appendix</TH>

@@

For distributional trees a prediction for a new set of covariates $\bold{z} = (z_1, \ldots, z_m)$ is made with the following strategy: 
All observations of the learning data set $\{(y_i,\bold{z}_i)\}_{i=1\ldots,n}$ which end up in the same terminal node of the tree as the new set of covariates are selected and a distributional model is fit on the subset. To select this subset a weight vector $\bold{w}^{\text{tree}} = (w^{\text{tree}}_1, \ldots, w^{\text{tree}}_n)$ is included in the fitting process.
$$
w^{\text{tree}}_i(\bold{z}) = \sum_{b=1}^B I((\bold{z}_i \in \mathcal{B}_b) \land (\bold{z} \in \mathcal{B}_b))
$$
The predicted distribution of the new set of covariates is then fully specified by the estimated parameter $\hat{\theta}(\bold{z})$ where
$$
\hat{\theta}(\bold{z}) = \max_{\theta \in \Theta} \sum_{i=1}^n w^{\text{tree}}_i(\bold{z}) \cdot \ell(\theta; y_i).
$$

While the clear structure of such a tree model is advantageous in interpreting and analyzing it, the abrupt changes are often too rough and impose steps on the model even if the true underlying effect is smooth. In these situations bagging can solve the problem by combining an ensemble of trees such that wrongly detected abrupt changes of the model are turned into almost smooth transitions. In that way, all kinds of effects can be modeled, which is is also one the of the main advantages of forest models (taking bagging one step further as the variance is reduced by sampling the covariates considered for the splitting process in each node).

%%% <TH> we need to be much more precise here: random forests is Breiman's and Cutler's
%%% random forest. Maybe it is better to not mention this method at any time. </TH>
Common forest models such as Breiman and Cutler's randomForest consist of trees only considering the mean of the distribution of the response variable. For that reason, changes in other parameters can only be detected if they are connected to changes in the location parameter, e.g. if they depend on the location parameter. Therefore, building a forest of distributional trees which can also capture separate effects in each of the distribution parameters leads to an improvement of the model. This valuable feature of detecting separate effects is also provided by GAMLSS as each parameter is modeled separately. However, in this case this separation can also be a disadvantage as it hinders the capturing of interactions unless they are specified explicitly by the user. These two cases provide once again a good reason to combine the two frameworks of flexible parametric modeling and tree-structured models in order to profit from their advantages and compensate possible disadvantages. This is the pursued goal in introducing distributional forests.

@@

Forest models differ mainly in the way how the trees are combined to make predictions. Following the strategy explained and proposed by \cite{Hothorn+Zeileis:2017} we interpret random forests as adaptive local likelihood estimators as distributional models are fit locally in subgroups while the term ``adaptive'' refers to the dependence on the splitting variables $\bold{Z}$ and the resulting subgroups and weighing schemes. In this case the forest itself is only used to evaluate neighborhood relationships among the observations which can then be included in the fitting process. In transformation forests by \cite{Hothorn+Zeileis:2017} so-called ``nearest neighbor weights'' are calculated as a measure of distance between observations. This approach has also already been applied in several random forest-type methods such as in ``adaptive nearest neighbors'' by \cite{Lin+Jeon:2006}, in ``quantile regression forests''  by  \cite{Meinshausen:2006} and in ``bagging survival trees'' by \cite{Hothorn+Lausen+Benner:2004}. The idea of nearest neighbor weights is to measure similarity of a pair of observations by counting the number of trees in which it is asigned to the same terminal node.
%Similar observations have a high probability of ending up in the same terminal node whereas this probability is low for quite different observations. 
%although the first random forest-type algorithm for the estimation of conditional distribution functions was published more than a decade ago (bagging survival trees, Hothorn et al. 2004). 
In particular, for a new set of covariates $\bold{z} = (z_1, \ldots, z_m) $ such a set of weights 
$\bold{w}^{\text{forest}}~=~(w^{\text{forest}}_1, \ldots, w^{\text{forest}}_n)$ is obtained by the following procedure: For each tuple $(y_i,\bold{z}_i)$ in the learning data set $\{(y_i,\bold{z}_i)\}_{i=1\ldots,n}$ a weight is calculated by counting the number of trees in which $(y_i,\bold{z}_i)$ ends up in the same terminal node as the new set of covariates $\bold{z}$:
$$
w^{\text{forest}}_i(\bold{z}) = \sum_{t=1}^T \sum_{b=1}^{B^t} I((\bold{z}_i \in \mathcal{B}^t_b) \land (\bold{z} \in \mathcal{B}^t_b))
$$
where $T$ is the number of trees of the forest, $B^t$ is the number of terminal nodes of the \textit{t}-th tree and $\mathcal{B}^t_b$ represents the \textit{b}-th terminal node of the \textit{t}-th tree.

Therefore, contrary to distributional trees, in distributional forests the whole learning data set can be used to fit the distributional model and specify the predicted distribution for a new set of covariates. Compared to the formula for predictions made by distributional trees the vector of tree weights $\bold{w}^{\text{tree}}$ containing only zeros and ones is now replaced by a vector of integer valued forest weights.
%%% <TH> we need references for the nearest neighbor weights, copy from
%%% "Transformation Forests" </TH>
%%%<TH> we need to cite "Transformation Forests" for the interpretation of
%%% random forests as adaptive local ML </TH>
This individual set of weights $\bold{w}^{\text{forest}}(\bold{z}) = (w^{\text{forest}}_1(\bold{z}), \ldots, w^{\text{forest}}_n(\bold{z}))$ for the new set of covariates $\bold{z}$ can now be included in the estimation process of the distribution parameters and in that way leads to its individual parameter vector $\hat{\theta}(\bold{z})$.
$$
\hat{\theta}(\bold{z}) =  \max_{\theta \in \Theta} \sum_{i=1}^n w^{\text{forest}}_i(\bold{z}) \cdot \ell(\theta; y_i)
$$
While in a distributional tree the same distribution is specified for all sets of covariates being in the same node, a distributional forest predicts and fits a different distribution for each different set of covariates due to the individual weights calculated separately for each set of covariates.

@@

%% TO DO: valueable features, why relevant in weather forecasting, what was missing here before? (complex terrain, many covariates, all kinds of effects could appear, interactions unknown)

@@

\item An ensemble model output statistics (EMOS, \citealp{Gneiting+Raftery+Westveld:2005}) model where the location parameter and the logarithm of the scale parameter depend linearly on the ensemble mean forecast of the total daily precipitation sum and the corresponding ensemble standard deviation respectively.
The choice of the logarithmic function as link function for the scale parameter has been made based on the results of \cite{Gebetsberger+Messner+Mayr:2017} where different link functions have been compared. Moreover, testing the identity, quadratic and logarithmic function as link functions in this setting supported this choice as the logarithmic function led to the best performance of the EMOS model in this application.

\item A GAMLSS where for each distribution parameter the selection of covariates and the specification of interactions is done in advance, both based on meteorological expert knowledge following \cite{Stauffer+Umlauf+Messner:2017} (prespecified GAM). Among the available 82 regressors, 9 where selected for the location parameter and 6 for the scale parameter. Additionally, for each distribution parameter one interaction term is specified in which convective available energy is interacting with total precipitation (with the ensemble mean for the location parameter and with the ensemble standard deviation for the scale parameter). 
%In this terms an ensemble mean of 6 hour sums of convective available energy is interacting with an ensemble mean or standarad deviation of 6 hour sums of total precipitation.  
These interactions might capture events with a high potential for moderate or severe thunderstorms with large amounts of precipitation. By putting extra effort in the specification of the formulas the prespecified GAM is provided with additional information and is therefore expected to profit from this advantage.
\item A boosted GAMLSS \citep{Hofner+Mayr+Schmid:2016} using non-homogeneous boosting
\citep{Messner+Mayr+Zeileis:2017} for an automatic selection of covariates (boosted GAM). The idea of this boosting approach is to iteratively minimize a loss function where in each step only the coefficient leading to the highest improvement of the current model is updated. As the initial coefficients are all zero, only the most important covariates are included with nonzero coefficients while less relevant variables are ignored if the procedure is stopped before convergence.
\end{itemize}

@@

\begin{table}[t!]
%\resizebox{1.4\textwidth}{!}{
%\begin{minipage}{\textwidth}
\centering
\begin{tabular}{ l l l l }
\hline
Model & Type & Location ($\mu$) & Scale ($\log(\sigma)$) \\
\hline
\textbf{Distributional forest} & recursive    & all & all \\ 
                & partitioning &     &     \\ 
\hline
\textbf{Prespecified GAM} & spline  & \emph{tp{\_}mean}, & \emph{tp{\_}sprd},\\
                            & in each & \emph{tp{\_}max},  & \emph{dswrf{\_}sprd{\_}mean},\\
 & & \emph{tp{\_}mean1218} $\ast$ & \emph{tp{\_}sprd1218} $\ast$\\ 
 & & \quad \emph{cape{\_}mean1218}, & \quad \emph{cape{\_}mean1218},\\
 & & \emph{dswrf{\_}mean{\_}mean}, & \emph{tcolc{\_}sprd{\_}mean},\\
 & & \emph{tcolc{\_}mean{\_}mean}, & \emph{tdiff500850{\_}mean}\\
 & & \emph{pwat{\_}mean{\_}mean}, & \\
 & & \emph{tdiff500850{\_}mean}, & \\
 & & \emph{msl{\_}diff} & \\
\hline
\textbf{Boosted GAM} & spline  & all & all \\
                       & in each &     &     \\
\hline
\textbf{EMOS} & linear & \emph{tp{\_}mean} & \emph{tp{\_}sprd} \\
\hline
\end{tabular}
\caption[Table caption text]{Overview of models with type of covariate dependency and included covariates for each distribution parameter.}
\label{model_specification}
%}
\end{table}

@@

\subsection{Application for one station}
\label{Application for one station}
First of all we focus on one observation station, Axams in Tyrol. This station was selected due to its geographical closeness to Innsbruck, the capital of Tyrol. As for all other stations, daily observations and numerical weather predictions are available for the month of July of 28 years from 1985 through the year 2012.
In Figure~\ref{axams_onestation} the results of fitting a distributional forest to the first 24 years of the data set and predicting the total amount of precipitation for one specific day of each of the four successive years have been shown as a motivational example for a probabilistic forecast. Staying with this fitted model we are now going to analyze variable importance of the distributional forest model and compare its results to those of the GAMs and the EMOS model based on residual QQ-plots.
%%% <TH> wouldn't be a uniform QQ-plot be more appropriate? </TH>

@@

%%% <TH> this should go into the discussion or the interpretation of the
%%% empirical results </TH>
%Even though many situations demand for the use of specifically tailored models providing the needed features and therefore being clearly advantageous and the best choice, it is often hard to tell in advance which features will be required. In these cases a distributional forest is a recommendable compromise providing reasonable models and performing well without any prespecification or additional knowledge needed in advance.

\section*{Computational details}
The corresponding implementation to the novel methods presented in this article is available on \textsf{R}-Forge in an \textsf{R} package called \textbf{disttree} within the project \textbf{partykit} (\url{https://R-Forge.R-project.org/projects/partykit/}). The function \code{disttree} uses either the \code{mob} or \code{ctree} function as a framework to build a tree. Similarly, the function \code{distforest} is based on the \code{cforest} function which provides a framework to build a distributional forest. All three functions, \code{mob}, \code{ctree} and \code{cforest}, are provided in the package \textbf{partykit} (\citealt{Hothorn+Zeileis:2015}). To specify a distribution family the user can hand over one of the following objects as input argument in \code{distfit}, \code{disttree} and \code{distforest}: 
\begin{itemize}
\item a GAMLLS family object from the \textsf{R} package \textbf{gamlss.dist} (\citealt{Stasinopoulos+Rigby:2007}).
\item a family list generating function from the package \textbf{disttree}.
\item a complete family list from the package \textbf{disttree}.
\item a custom list containing all required information about the distribution family.
\end{itemize}

@@

%\newpage
\nocite{Seibold+Zeileis+Hothorn:2016, Seibold+Zeileis+Hothorn:2017, Yee:2010}
% Zeileis+Wiel+Hornik:2008, 
% Zeileis+Hothorn+Hornik:2010,

>>>>

\subsection{Motivation}
As illustrated in Figure~\ref{Devel_parmod} a wide range of tools has been developed 
in the field of parametric modeling. Especially the shift from conditional mean 
regression models to distributional regression models has increased the application 
spectrum by allowing for much broader inference. In some cases further features 
such as the ability to capture non-additive effects or interactions are desirable 
which demands for more flexibility. Since tree-structured models provide these 
required features, combining them with parametric models can offer a framework 
that is able to deal with a wide variety of different modeling situations. 
In this section the novel methodology which embeds parametric modeling into 
the idea of tree-structured models will be introduced step by step. Starting 
with a simple distributional fit in Section~\ref{distfit} will be followed by 
building a distributional tree in Section~\ref{disttree} and finally a distributional 
forest in Section~\ref{distforest}.

@@

In our application of precipitation forecasting the response variable, which is 
the power transformed total precipitation amount within 24 hours, is expected 
to be normally distributed with mean/location parameter $\mu$ and standard 
deviation/scale parameter $\sigma$ and left-censored with censoring point 0. 
(The power transformation is used to remove large parts of the skewness of 
the data as further explained in Section~\ref{data}.) Therefore, the 
corresponding log-likelihood function with parameter vector 
$\theta = (\mu, \sigma)$ is
$$
\ell(\mu, \sigma; Y) = 
\begin{cases}
    \log\left(\frac{1}{\sigma} \cdot \phi\left(\frac{Y - \mu}{\sigma}\right) \right), & \text{if } Y > 0\\
    \vspace{0.1cm}
    \log\left(\Phi\left(\frac{-\mu}{\sigma}\right)\right), & \text{if } Y = 0
\end{cases}
$$
where $\phi$ and $\Phi$ are the probability density function and the 
distribution function of the standard normal distribution $\mathcal{N}(0,1)$. 
Next to the censored normal distribution also the censored logistic 
distribution is often chosen for the modeling of precipitation amounts. In 
\cite{Gebetsberger+Messner+Mayr:2017} this choice has been made in order 
to be able to deal with heavy tails which often appear in precipitation 
data. Another frequent application of probabilistic forecasting in meteorology 
is the modeling of near surface air temperature. In this case a normal 
distribution is often selected.

With the specification of the distribution family and its log-likelihood 
function the task of fitting a distributional model turns into the task 
of determining the distribution parameter~$\theta$. This is done by the 
following maximization.
$$
\hat{\theta} = \max_{\theta \in \Theta} \sum_{i=1}^n \ell(\theta; y_i)
$$
where $\{y_i\}_{i=1,\ldots,n}$ are the observations of the response 
variable $Y$.

In that way, a whole distribution is specified with all its features, 
including location, scale, and shape. However, fitting one global model 
to the whole data set might be too much of a generalization and therefore 
not represent the data and its features in a reasonable way. Moreover, 
if covariates are available it is desirable to include them into the model 
as they might provide important additional information. In our application 
on precipitation data a wide range of covariates is available from a numerical 
ensemble prediction system considering precipitation, temperature, sea level 
pressure and other meteorological quantities which might be of interest and 
have an influence on the predictions.

@@

\subsection{Distributional tree}
\label{disttree}
Given the fit of a global distributional model $D(Y, \bold{\hat{\theta}})$ 
the idea is to use the information provided by this model together with a 
set of covariates $\bold{Z} = (Z_1, \ldots, Z_m) \in \mathcal{Z}$ 
to decide whether the data set should be split into subgroups or not. 
In particular, the information provided by the fitted model are the scores 
$$
s(\hat{\theta}, y_i) = \frac{\partial \ell}{\partial \theta}(\hat{\theta}; y_i).
$$ 
These values express how well the model fits the data regarding each 
parameter. Ideally they should fluctuate randomly around zero, 
similar to residuals in ordinary least squares estimation, but in 
this case there is a score value for each pair of estimated parameter 
$\hat{\theta}_j \in (\hat{\theta}_1, \ldots, \hat{\theta}_k) = \bold{\hat{\theta}}$ 
and observation $y_j \in \{y_i\}_{i = 1,\ldots, n}$. This enables the 
model to be sensitive to changes in each of the parameters which is one 
of the main advantages of distributional trees over other tree models 
considering only the mean of the distribution.

Statistical tests are applied to check whether there is a significant 
dependency between the scores and each of the partitioning variables 
$Z_l \in \bold{Z} = (Z_1, \ldots, Z_m)$. If so, the data set is split 
into subgroups based on the values of the partitioning variable that 
shows the highest dependency. The exact way of determining how and where 
to split varies across different tree algorithms. For distributional trees 
one possible choice is a permutation test based algorithm which is explained 
in detail in the appendix.

After splitting the data the next step is to 
fit a local distributional model in each of the subgroups. Repeating this 
procedure of fitting a model, applying statistical tests to evaluate parameter 
instability and splitting the data set depending on the test results leads 
to a tree structure. In that way the learning data set is separated into 
\textit{B} disjoint segments such that
$$
\mathcal{Z} = \dot{\bigcup\limits_{b = 1, \ldots, \textit{B}}} \mathcal{B}_b
$$
where each segment is represented by a terminal node of the tree.

Summed up, the steps of building a distributional tree are:
\begin{enumerate}
\item Specify a distribution family with log-likelihood function $\ell(\theta; y)$.
\item Estimate $\bold{\hat{\theta}}$ by solving $\sum_{i=1}^n s(\hat{\theta}, y_i) = 0$.
\item Test for associations or instabilities of the scores $s(\bold{\hat{\theta}}, y_i)$ 
and each partitioning variable~$Z_l$.
\item Split the sample along the partitioning variable with the strongest association 
or instability. Choose the breakpoint with the highest improvement in log-likelihood.
\item Repeat steps 2--4 recursively in the subgroups until some stopping criterion is reached.

\end{enumerate}
The stopping criterion in step 5 can for example be a minimum number of 
observations in a node or a p-value for the statistical tests applied in step 3.
%%% <TH> we need a little more detail here or a reference to a better
%%% description of mob/ctree OR appendix</TH>

@@

For distributional trees a prediction for a (possibly new) set of covariates 
$\bold{z} = (z_1, \ldots, z_m)$ is made with the following strategy: 
All observations of the learning data set $\{(y_i,\bold{z}_i)\}_{i=1\ldots,n}$ 
which end up in the same terminal node of the tree as $\bold{z}$ are selected 
and a distributional model is fit on this subset. Selecting these observations
is done by including a weight vector 
$\bold{w}^{\text{tree}} = (w^{\text{tree}}_1, \ldots, w^{\text{tree}}_n)$ 
in the fitting process where
$$
w^{\text{tree}}_i(\bold{z}) = \sum_{b=1}^B I((\bold{z}_i \in \mathcal{B}_b) \land (\bold{z} \in \mathcal{B}_b)).
$$
The predicted distribution for $\bold{z}$ is then fully 
specified by the estimated parameter $\hat{\theta}(\bold{z})$ where
$$
\hat{\theta}(\bold{z}) = \max_{\theta \in \Theta} \sum_{i=1}^n w^{\text{tree}}_i(\bold{z}) \cdot \ell(\theta; y_i).
$$

While the clear structure of such a tree model is advantageous in 
interpreting and analyzing it, the abrupt changes are often too rough 
and impose steps on the model even if the true underlying effect is 
smooth. In these situations bagging can solve the problem by combining 
an ensemble of trees such that wrongly detected abrupt changes of the 
model are turned into almost smooth transitions. In that way, all kinds 
of effects can be modeled, which is also one the of the main advantages 
of forest models taking bagging one step further as the variance is 
reduced by sampling the covariates considered for the splitting process 
in each node.

%%% <TH> we need to be much more precise here: random forests is Breiman's and Cutler's
%%% random forest. Maybe it is better to not mention this method at any time. </TH>
Common forest models (such as Breiman and Cutler's randomForest) consist of 
trees only considering the mean of the distribution of the response variable. 
For that reason, changes in other parameters can only be detected if they 
are connected to changes in the location parameter, e.g. if they depend on 
the location parameter. As explained before, a distributional tree considers
scores for all distribution parameters. Therefore, a forest consisting of 
distributional trees can also capture separate effects in each of the 
distribution parameters. For that reason distributional forests are
now going to be introduced.

@@

Forest models differ mainly in the way how the trees are combined to make predictions. 
Following the strategy explained and proposed by \cite{Hothorn+Zeileis:2017} we 
interpret random forests as adaptive local likelihood estimators as distributional 
models are fit locally in subgroups while the term ``adaptive'' refers to the 
dependence on the splitting variables $\bold{Z}$ and the resulting subgroups 
and weighing schemes. In this case the forest itself is only used to evaluate 
neighborhood relationships among the observations which can then be included 
in the fitting process. In transformation forests by \cite{Hothorn+Zeileis:2017} 
so-called ``nearest neighbor weights'' are calculated as a measure of distance 
between observations. This approach has also already been applied in several 
random forest-type methods such as in ``adaptive nearest neighbors'' by 
\cite{Lin+Jeon:2006}, in ``quantile regression forests''  by  \cite{Meinshausen:2006} 
and in ``bagging survival trees'' by \cite{Hothorn+Lausen+Benner:2004}. 
The idea of nearest neighbor weights is to measure similarity of a pair of 
observations by counting the number of trees in which it is asigned to the 
same terminal node.
%Similar observations have a high probability of ending up in the same terminal 
%node whereas this probability is low for quite different observations. 
%although the first random forest-type algorithm for the estimation of 
%conditional distribution functions was published more than a decade ago 
%(bagging survival trees, Hothorn et al. 2004). 
In particular, for a (possibly new) set of covariates 
$\bold{z} = (z_1, \ldots, z_m) $  such a set of weights 
$\bold{w}^{\text{forest}}~=~(w^{\text{forest}}_1, \ldots, w^{\text{forest}}_n)$ 
is obtained by the following procedure: For each tuple $(y_i,\bold{z}_i)$ in 
the learning data set $\{(y_i,\bold{z}_i)\}_{i=1\ldots,n}$ a weight is calculated 
by counting the number of trees in which $(y_i,\bold{z}_i)$ ends up in the same 
terminal node as $\bold{z}$:
$$
w^{\text{forest}}_i(\bold{z}) = \sum_{t=1}^T \sum_{b=1}^{B^t} I((\bold{z}_i \in \mathcal{B}^t_b) \land (\bold{z} \in \mathcal{B}^t_b))
$$
where $T$ is the number of trees of the forest, $B^t$ is the number of terminal 
nodes of the \textit{t}-th tree and $\mathcal{B}^t_b$ represents the \textit{b}-th 
terminal node of the \textit{t}-th tree.

Therefore, contrary to distributional trees, in distributional forests the whole 
learning data set can be used to fit a distributional model and specify the 
predicted distribution for a set of covariates. Compared to the formula for
predictions made by distributional trees the vector of tree weights 
$\bold{w}^{\text{tree}}$ containing only zeros and ones is now replaced by a 
vector of integer valued forest weights.
%%% <TH> we need references for the nearest neighbor weights, copy from
%%% "Transformation Forests" </TH>
%%%<TH> we need to cite "Transformation Forests" for the interpretation of
%%% random forests as adaptive local ML </TH>
This individual set of weights 
$\bold{w}^{\text{forest}}(\bold{z}) = (w^{\text{forest}}_1(\bold{z}), \ldots, w^{\text{forest}}_n(\bold{z}))$ 
for the set of covariates $\bold{z}$ can now be included in the 
estimation process of the distribution parameters and in that way leads 
to its individual parameter vector $\hat{\theta}(\bold{z})$ where
$$
\hat{\theta}(\bold{z}) =  \max_{\theta \in \Theta} \sum_{i=1}^n w^{\text{forest}}_i(\bold{z}) \cdot \ell(\theta; y_i)
$$
While in a distributional tree the same distribution is specified for all 
sets of covariates being in the same node, a distributional forest predicts 
and fits a different distribution for each different set of covariates due 
to the individual weights calculated separately for each set of covariates.

@@

@@

\item An ensemble model output statistics 
(EMOS, \citealp{Gneiting+Raftery+Westveld:2005}) model where the 
location parameter and the logarithm of the scale parameter depend 
linearly on the ensemble mean forecast of the total daily precipitation 
sum and the corresponding ensemble standard deviation respectively.
The choice of the logarithmic function as link function for the scale 
parameter has been made based on the results of 
\cite{Gebetsberger+Messner+Mayr:2017} where different link functions 
have been compared. Moreover, testing the identity, quadratic and 
logarithmic function as link functions in this setting supported 
this choice as the logarithmic function led to the best performance 
of the EMOS model in this application.
\item A GAMLSS where for each distribution parameter the selection 
of covariates and the specification of interactions is done in 
advance, both based on meteorological expert knowledge following 
\cite{Stauffer+Umlauf+Messner:2017} (prespecified GAM). Among the 
available 82 regressors, 9 where selected for the location parameter 
and 6 for the scale parameter. Additionally, for each distribution 
parameter one interaction term is specified in which convective 
available energy is interacting with total precipitation (with the 
ensemble mean for the location parameter and with the ensemble 
standard deviation for the scale parameter). These interactions 
might capture events with a high potential for moderate or severe 
thunderstorms with large amounts of precipitation. By putting 
extra effort in the specification of the formulas the prespecified 
GAM is provided with additional information and is therefore 
expected to profit from this advantage.
\item A boosted GAMLSS \citep{Hofner+Mayr+Schmid:2016} using 
non-homogeneous boosting \citep{Messner+Mayr+Zeileis:2017} 
for an automatic selection of covariates (boosted GAM). 
The idea of this boosting approach is to iteratively minimize 
a loss function where in each step only the coefficient leading 
to the highest improvement of the current model is updated. As 
the initial coefficients are all zero, only the most important 
covariates are included with nonzero coefficients while less 
relevant variables are ignored if the procedure is stopped before 
convergence.
\end{itemize}

@@

\begin{table}[t!]
\centering
\begin{tabular}{ l l l l }
\hline
Model & Type & Location ($\mu$) & Scale ($\log(\sigma)$) \\
\hline
\textbf{Distributional forest} & recursive    & all & all \\ 
                & partitioning &     &     \\ 
\hline
\textbf{Prespecified GAM} & spline  & \emph{tp{\_}mean}, & \emph{tp{\_}sprd},\\
                            & in each & \emph{tp{\_}max},  & \emph{dswrf{\_}sprd{\_}mean},\\
 & & \emph{tp{\_}mean1218} $\ast$ & \emph{tp{\_}sprd1218} $\ast$\\ 
 & & \quad \emph{cape{\_}mean1218}, & \quad \emph{cape{\_}mean1218},\\
 & & \emph{dswrf{\_}mean{\_}mean}, & \emph{tcolc{\_}sprd{\_}mean},\\
 & & \emph{tcolc{\_}mean{\_}mean}, & \emph{tdiff500850{\_}mean}\\
 & & \emph{pwat{\_}mean{\_}mean}, & \\
 & & \emph{tdiff500850{\_}mean}, & \\
 & & \emph{msl{\_}diff} & \\
\hline
\textbf{Boosted GAM} & spline  & all & all \\
                       & in each &     &     \\
\hline
\textbf{EMOS} & linear & \emph{tp{\_}mean} & \emph{tp{\_}sprd} \\
\hline
\end{tabular}
\caption[Table caption text]{Overview of models with type of 
covariate dependency and included covariates for each distribution 
parameter.}
\label{model_specification}
\end{table}

@@

\subsection{Application for one station}
\label{Application for one station}
First of all we focus on one observation station, Axams in Tyrol. This 
station was selected due to its geographical closeness to Innsbruck, 
the capital of Tyrol. As for all other stations, daily observations and 
numerical weather predictions are available for the month of July of 
28~years from 1985 through the year 2012. In Figure~\ref{axams_onestation} 
the results of fitting a distributional forest to the first 24 years of the 
data set and predicting the total amount of precipitation for one specific 
day of each of the four successive years have been shown as a motivational 
example for a probabilistic forecast. Staying with this fitted model we are 
now going to analyze variable importance of the distributional forest model 
and compare its results to those of the GAMs and the EMOS model based on 
residual QQ-plots.

@@

%Even though many situations demand for the use of specifically tailored 
%models providing the needed features and therefore being clearly advantageous 
%and the best choice, it is often hard to tell in advance which features 
%will be required. In these cases a distributional forest is a 
%recommendable compromise providing reasonable models and performing 
%well without any prespecification or additional knowledge needed in advance.

\section*{Computational details}
The corresponding implementation to the novel methods presented 
in this article is available on \textsf{R}-Forge in an \textsf{R} 
package called \textbf{disttree} within the project \textbf{partykit} 
(\url{https://R-Forge.R-project.org/projects/partykit/}). The function 
\code{disttree} uses either the \code{mob} or \code{ctree} function 
as a framework to build a tree. Similarly, the function 
\code{distforest} is based on the \code{cforest} function which 
provides a framework to build a distributional forest. All three 
functions, \code{mob}, \code{ctree} and \code{cforest}, are provided 
in the package \textbf{partykit} (\citealt{Hothorn+Zeileis:2015}). 
To specify a distribution family the user can hand over one of the 
following objects as input argument in \code{distfit}, \code{disttree} 
and \code{distforest}: 
\begin{itemize}
\item a GAMLLS family object from the \textsf{R} package \textbf{gamlss} 
(\citealt{Stasinopoulos+Rigby:2007}).
\item a family list generating function from the package \textbf{disttree}.
\item a complete family list from the package \textbf{disttree}.
\item a custom list containing all required information about the 
distribution family.
\end{itemize}

@@

\nocite{Seibold+Zeileis+Hothorn:2016, Seibold+Zeileis+Hothorn:2017, Yee:2010}
% Zeileis+Wiel+Hornik:2008, 
% Zeileis+Hothorn+Hornik:2010,
